Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2017-08-10 paper.tex
New figure 1
New figure 1 version 2 =)
More edits from Kang-Kuen

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite  a  large Lamb-Dicke parameter, high initial temperature, and
  large differential light shifts between the excited state and the ground state,
  we achieve a ground state population of $81(4)$\% after $100$ ms of cooling. In addition, $10\,\%$ of initial atom population escapes the trap during cooling.
  Our technique includes addressing  high-order sidebands, , where several motional quanta are removed by a single laser pulse,  and
  fast modulation of the optical tweezer intensity.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is possible %outside the Lamb-Dicke regime, 
  where tight confinement and low initial temperature are difficult to realize.
%  This is particularly relevant for systems which are challenging to laser-cool, such as molecules and exotic atoms, and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

Extending optical tweezer arrays  to include polar molecules will open up a  range of
new applications that exploit  long-lived molecular internal states
and  tunable long range interactions~\cite{DeMille2002,Gorshkov2011,Yan2013}.
Molecules could be assembled from atom pairs in  optical tweezers~\cite{Liu2017},
or directly loaded from magneto-optical traps
(MOTs)~\cite{Barry2014,Truppe2017SubDoppler,Anderegg2017}.
For either approach, preparing the constituent atoms or molecules in the lowest motional
quantum state is important to realize long coherence times in quantum applications.

Preparation of single atoms in the motional ground state has been achieved in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
 using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems take place in the Lamb-Dicke (LD) regime, where
the spread of the initial wavefunctions is much smaller than the wavelength of light used to address them.
Systems with a large wavefunction, either due to small mass or high initial temperature,
fall outside of the LD regime and undergo strong recoil heating during RSC.
Extending RSC beyond the LD regime opens  the possibility
for ground state cooling of systems such as light atoms and %systems with poor cooling such as 
molecules that can be laser-cooled.
In this letter we demonstrate cooling of single sodium atoms in optical tweezers,
initially outside of the LD regime, to the motional ground state.
We achieve a 3D ground state population of $P_0=81(4)$\% by  cooling via
 high-order sidebands in a carefully optimized cooling sequence.
Our approach is general and applicable to %opens up ground state cooling for 
other systems.

Our experiment has an overall repetition rate of $2.5$ Hz and
begins by loading a single sodium atom into optical tweezers from a MOT~\cite{Hutzler2017-LightShifts}.
%for $250$ ms~\cite{Hutzler2017-LightShifts}.
The tweezers are created by focussing a $700$ nm wavelength laser beam through an NA=$0.55$ objective to an elliptical waist with radii $\{w_{0,x},w_{0,y}\}\approx \{0.79,0.57\}\,\mu$m.
For $45$ mW of optical power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{430(4), 590(5), 69(1)\}\ \text{\text{kHz}}$,
where  $z$ is the weakly confined axial direction,
and  $x$ and $y$ are the tightly confined radial directions.
After each single atom loading attempt, an image is taken with a $1.5$ ms exposure time
to determine success.
During  imaging, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to mean motional states along the three trap axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.4(2),\, 2.5(2),\, 21(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
The atom is then initialized  into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping (OP).
Each experimental cycle ends with a second imaging sequence that measures the hyperfine state of the atom as $|F=2\rangle$ or $|F=1\rangle$~\footnote{The final hyperfine state sensitive detection employs a strong beam that is resonant with the $F=2$ to $F'=3$ cycling transition, to heat the $F=2$ population out of the trap.  Therefore, any atoms which survive this heat out procedure are interpreted as having been in the $F=1$ state.}.%a repetition of the experiment that was initially bright and remained bright after the pulse sequence is interpreted as a final F=1 atomic state.}.

@@

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{Single Na atom Raman sideband cooling scheme and sequence. (A)
    The Raman transitions between $|2,-2;n\rangle$ and $|1,-1;n+\Delta n\rangle$ have a one-photon detuning $\Delta=25$ GHz below the $3^2S_{1/2}$ to $3^2P_{3/2}$ transition. Two-photon detuning, $\delta$, is defined relative to the $\Delta n=0$ carrier transition. For optical pumping, we use two $\sigma^-$ polarized transitions, one   to  pump the atom state out of $|1,-1\rangle$ via $3^2P_{3/2}$ and one  to pump atoms out of $|2,-1\rangle$ via $3^2P_{1/2}$
     to minimize heating on the atom in the $|2,-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezers and bias magnetic field.  Raman beams 1 and 4 address the radial $x$ mode. Raman beams 1 and 2 address the radial $y$ mode. Raman beams 3 and 4 address the axial $z$ mode, which also couples to the radial $y$ mode.
    (C) Schematic of the cooling pulse sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ sideband pulses.
    The four axial pulses address two sideband orders.
    The two pulses in each radial direction  either address $\Delta n=-2$ and $\Delta n=-1$
    or have different durations to drive $\Delta n=-1$, when most of the population is below $n=3$
    at the end of the cooling sequence.
    The Raman cooling and spectroscopy pulses have Blackman  envelopes to reduce  off-resonant
    coupling, while the measurement Rabi pulses in Fig.~\ref{f-radial}(B,C) and \ref{f-axial}(B,C) have square envelopes to simplify analysis.
    \label{f-setup}}
\end{figure*}

Specifically, Raman transitions are driven in Na between the hyperfine states
$|2, -2\rangle$ and $|1, -1\rangle$ in the presence of an $8.8$ G magnetic field. The external field is
orthogonal to the effective magnetic field of the tweezer, to reduce  vector light shifts~\cite{Kaufman2012,Thompson2013}.
Subsequently, an OP pulse, which consists of two frequencies  both with $\sigma^-$-polarization, is applied  to bring the atom out of the $F=1$ manifold and back to  $|2, -2\rangle$ via spontaneous emission (Fig.~\ref{f-setup}A).
%brings the atom back to $|F=2, m_F=-2\rangle$ via spontaneous emission.
%, which consists of two frequencies with $\sigma^-$-polarization for the  to bring the atom out of $F=1$ manifold and back to  $|F=2, m_F=-2\rangle$ via spontaneous emission.
It is important that the optical pumping step scatter as few photons as possible to minimize recoil heating. %The OP consists of driving a $|F=1, m_F=-1\rangle$ into the $F=2$ manifold via the transition of $3P_{3/2} D2 line$
Therefore, to optically pump atoms from $|2,-1\rangle$ into $|2, -2\rangle$, we apply light resonant with  $3^2P_{1/2}$~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3^2P_{3/2}$, from which the $|2, -2>$ state could always scatter a photon via the excited $|F'=3, m_F'=-3>$ state.} because $|2, -2\rangle$ is a dark state to $\sigma^-$-polarized light~\cite{Monroe1995, Grobner2017}.
%~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3P_{3/2}$, from which the $|2, -2 \rangle$}

\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping motional-state redistribution and  Raman coupling for large  Lamb-Dicke parameters for the axial direction ($z$). The range plotted covers $95$\% of the initial thermal distribution.
   % The range plotted covers $95$\% of the initial thermal distribution.
    (A) Motional state distribution after one optical pumping cycle
    for different initial n, $n_{\textrm{init}}$. Due to photon-recoil and the large Lamb-Dicke parameter, $\eta^{OP}_z=0.6$,
    there is a high  $n$ changing probability.         (B) Matrix elements for Raman transition deviate from
    $\sqrt{n}$ scaling with multiple minima. During cooling, we utilize the fact that higher motional states couple most effectively to sidebands with large $|\Delta n|$.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in the axial direction).
    \label{f-ld}}
\end{figure}

For an atom starting in motional level $n_{\textrm{init}}$, the probability of recoil during the OP sequence results in
a motional-state redistribution  is shown in Fig.~\ref{f-ld}A.
The  probability is approximately proportional to the square of effective LD parameter
$\eta^{\textrm{OP}}_{\textrm{eff}}=\sqrt{2n+1}\eta^{\textrm{OP}}$~\cite{Kaufman2012},
where the optical pumping LD parameter $\eta^{\textrm{OP}} = z_0\frac{2\pi}{\lambda}$ is the ratio of the zero-point wavefunction spread $z_0 = \sqrt{\hbar/2m\omega}$ to the reduced wavelength $\lambda$.
Here, $m$ is the mass of the atom, $\omega$ is the trap frequency,
and $\hbar$ is the reduced Planck constant.
Such heating presents a major challenge for efficient cooling of single Na atoms
which feature a low mass (23 amu) and short wavelengths.
For our configuration, 
$\{\eta^{\textrm{OP}}_x,\eta^{\textrm{OP}}_y,\eta^{\textrm{OP}}_z\} = \{0.241(2), 0.206(1), 0.602(5)\}$.
A relatively large PGC temperature results in initial $\{\eta^{\textrm{OP}}_{x,\textrm{eff}},\eta^{\textrm{OP}}_{y,\textrm{eff}},\eta^{\textrm{OP}}_{z,\textrm{eff}}\} = \{0.67(2), 0.50(1), 4.0(1)\}$.
The weak axial direction %, $(\eta^{\textrm{OP}}_{z,\textrm{eff}})^2>1$ 
is far outside of the LD regime ($\eta^2\ll 1$), and the probability of recoil heating per OP step is large. %~\footnote{For example, the average change in the motional states for the initial distribution is $2.4$ in the axial direction, which is significantly greater than $0.89-0.95$in previous experiments of RSC of Rb and Cs~\cite{Li2012,Kaufman2012,Thompson2013,Liu2017}.},
This creates a large uncertainty on the motional state at the end of a full RSC step, possibly leading to runaway heating.

Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$,
where the wave vector is replaced by the wave vector difference $\Delta k$
between the two beams that drive the Raman transition.
The Raman transitions in our configuration have LD parameters of
$\{\eta^R_x,\eta^R_y,\eta^R_z\} = \{0.341(2), 0.291(1), 0.40(1)\}$.
To offset heating from OP initially, higher-order Raman  sidebands ($\Delta n <-1$)  can be utilized
to remove several motional quanta in a single cooling pulse (see Fig.~\ref{f-ld}B).
Because the coupling strengths of different orders do not reach minima for the same state of motion $n$,
using multiple orders of sidebands avoids accumulations of population
near the coupling minima.

Taking the large motional-state changing heating and cooling sources
 into account,
it is not immediately clear that efficient cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search~\cite{Dalibard1992}. 
%and to find a robust cooling sequence~\footnote{See supplemental material}.
In particular, we find that alternating the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and $\Delta n=-2$ and $\Delta n=-1$ for the radial directions
eliminates the accumulation of population in motional states that have zero Raman coupling, which would halt the cooling process.
The simulation also indicates that setting the coupling strength of each  sideband
to drive a Rabi $\pi$-pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in Fig.~\ref{f-ld}B)  yields efficient cooling, initially.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state, so the final  cycles utilize only the $\Delta n=-1$ sideband while alternating between the three  axes.

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed  sideband orders ($\Delta n=-8$ and $\Delta n=-7$)
and decreasing the orders to the next pair after every $6$ to $15$ cycles.
This process is repeated until we reach $\Delta n=-1$, and most of the
population is in the first few excited states. We then switch to cooling only on $\Delta n=-1$ with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
Radial cooling is performed similarly to the axial cooling with the initial cooling of $\Delta n=-2$ and $\Delta n=-1$ and switching to $\Delta n=-1$ only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which can then be used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.

There are two additional important challenges in cooling single sodium atoms.
First, the high initial temperature populates high motional states,
causing the atoms to sample the anharmonicity of the trap away from the center. Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.  We find $A_{i,n}=\frac{3n \hbar}{4 \pi m d_i^2}$  where $d_i=w_{0,x},w_{0,y}$ for the radial directions and $d_z=\pi w_{0,x}w_{0,y}/\lambda$.  Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{1.1, 2.0, 0.16\}n$~kHz. For mixed states, this  broadens high-order sidebands due to the $n$-dependence of the transitions. 
In order to mitigate this, we drive the sidebands with a large  Rabi frequency
and short pulses to Fourier broaden the spectrum. %, and use Blackman pulse envelopes to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}. 
However, anharmonicity  may still shift high-$n$ states out of resonance and prevent effective cooling.
%Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.  We find $A_{i,n}=\frac{3n \hbar}{2 m d_i^2}$  where $d_i=w_0$ for the radial directions and $d_z=\pi w_0^2/\lambda$  Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{1.3, 1.2, 0.2\}n$~kHz.  Anharmonicity broadens the high-order axial transitions and may shift high-$n$ states out of resonance to prevent effective cooling.

@@

We measure a heating rate that corresponds to  decreasing of 3D ground state population of single atoms in the tweezers  of  $\sim0.9$\%/ms. 
The  rate is consistent with off-resonant scattering of the trapping light~\cite{Grimm2000}, and is predominantly in the axial direction where the trapping beam propagates. 
%The ground state probability achieved is currently limited by off-resonant scattering from the Raman beams, which are measured to be between $3$ to $15$ kHz for the detuning and power we use, and the resonance shift caused by magnetic field fluctuations.

>>>>

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite  a  large Lamb-Dicke parameter, high initial temperature, and
  large differential light shifts between the excited state and the ground state,
  we achieve a ground state population of $81(4)$\% after $100$ ms of cooling, for the 85\% of atoms that survive cooling and re-imaging.
  Our technique includes addressing  high-order sidebands, where several motional quanta are removed by a single laser pulse,  and
  fast modulation of the optical tweezer intensity.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is possible, even without  tight confinement and low initial temperature.
%  This is particularly relevant for systems which are challenging to laser-cool, such as molecules and exotic atoms, and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

Extending optical tweezer arrays  to include polar molecules will open  a  range of
new applications that exploit  long-lived molecular internal states
and  tunable long range interactions~\cite{DeMille2002,Gorshkov2011,Yan2013}.
Molecules could be assembled from atom pairs in  optical tweezers~\cite{Liu2017},
or directly loaded from magneto-optical traps
(MOTs)~\cite{Barry2014,Truppe2017SubDoppler,Anderegg2017}.
For either approach, preparing the constituent atoms or molecules in the lowest motional
quantum state is important to realize long coherence times in quantum applications.

Preparation of single atoms in the motional ground state has been achieved in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
 using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems takes place in the Lamb-Dicke (LD) regime, where
the spread of the initial wavefunctions is  smaller than the reduced wavelength of light ($\lambdabar$)
used to address them.
Systems with a large RMS position spread ($z_{rms}$), either due to small mass or high initial temperature,
fall outside of the LD regime and undergo strong recoil heating during RSC.
In this letter we demonstrate cooling of single sodium atoms in optical tweezers,
initially outside  the LD regime ($z_{rms}/\lambdabar=4.0$), to the motional ground state.
We achieve a 3D ground state population of $P_0=81(4)$\% by  cooling via
 high-order sidebands in a carefully optimized cooling sequence.
Our approach is general and applicable to %opens up ground state cooling for
other systems.  Extending RSC beyond the LD regime opens  the possibility
for ground state cooling of systems such as light atoms and %systems with poor cooling such as
molecules that can be laser-cooled.

Our experiment has an overall repetition rate of $2.5$ Hz and
begins by loading a single sodium atom into an optical tweezer from a MOT~\cite{Hutzler2017-LightShifts}.
%for $250$ ms~\cite{Hutzler2017-LightShifts}.
The tweezer is created by focussing a $700$ nm wavelength laser beam through an NA=$0.55$ objective to an elliptical waist with radii $\{w_{0,x},w_{0,y}\}\approx \{0.79,0.57\}\,\mu$m.
For $45$ mW of optical power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{430(4), 590(5), 69(1)\}\ \text{\text{kHz}}$,
where  $z$ is the weakly confined axial direction,
and  $x$ and $y$ are the tightly confined radial directions.
After each single atom loading attempt, a first image is taken with a $1.5$ ms exposure time
to determine success, which occurs about 50\% of the time due to the loading mechanism~\cite{Schlosser2001}.
%After attempting to load a single atom from the MOT, an image with 1.5 ms exposure time is taken to determine if the attempt was successful, which occurs about 50% of the time due to the loading mechanism [ref].
During  imaging, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to mean motional states along the three trap axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.4(2),\, 2.5(2),\, 21(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
The atom is then initialized  into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping (OP).
Each experimental cycle ends with a second imaging sequence that measures the hyperfine state of the atom as $|F=2\rangle$ or $|F=1\rangle$~\footnote{The final hyperfine state sensitive detection employs a strong beam that is resonant with the $F=2$ to $F'=3$ cycling transition, to heat the $F=2$ population out of the trap.  Therefore, any atoms which survive this heat out procedure are interpreted as having been in the $F=1$ state.}. The data analysis includes only those experiments where the first image  reveals successful atom loading.
%a repetition of the experiment that was initially bright and remained bright after the pulse sequence is interpreted as a final F=1 atomic state.}.

@@

\begin{figure*}
  \includegraphics[height=4.5cm]{fig1_combined.pdf}
  \caption{Single Na atom Raman sideband cooling scheme and sequence. (A)
    The Raman transitions between $|2,-2;n\rangle$ and $|1,-1;n+\Delta n\rangle$ have a one-photon detuning $\Delta=25$ GHz below the $3^2S_{1/2}$ to $3^2P_{3/2}$ transition. Two-photon detuning, $\delta$, is defined relative to the $\Delta n=0$ carrier transition. For optical pumping, we use two $\sigma^-$ polarized transitions, one   to  pump the atom state out of $|1,-1\rangle$ via $3^2P_{3/2}$ and one  to pump atoms out of $|2,-1\rangle$ via $3^2P_{1/2}$
     to minimize heating of the $|2,-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.  Raman beams R1 and R4 address the radial $x$-mode. R1 and R2 address the radial $y$-mode.  R3 and R4 address the axial $z$-mode, where the beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.  %, which also couples to the radial $y$ mode.    The radial $x$-mode is addressed by turning on beams R1 and R4, $y$ is addressed by R1 and R2, and $z$ is addressed by R3 and R4.  In this geometry, the radial mode directions coincide with the respective $k$-vector differences.  The ``axial'' Raman beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.
    (C) Schematic of the cooling pulse sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ sideband pulses.
    The four axial pulses address two sideband orders.
    The two pulses in each radial direction  either address $\Delta n=-2$ and $\Delta n=-1$
    or have different durations to drive $\Delta n=-1$, at the end of the cooling sequence when most of the population is below $n=3$.
    The Raman cooling and spectroscopy pulses have Blackman  envelopes to reduce  off-resonant
    coupling, while the measurement Rabi pulses in Fig.~\ref{f-radial}(B,C) and \ref{f-axial}(B,C) have square envelopes to simplify analysis.
    \label{f-setup}}
\end{figure*}

Specifically, Raman transitions are driven in $^{23}$Na between the hyperfine states
$|2, -2\rangle$ and $|1, -1\rangle$ in the presence of an $8.8$ G magnetic field. The external field is
orthogonal to the effective magnetic field of the tweezer, to reduce  vector light shifts~\cite{Kaufman2012,Thompson2013}.
Subsequently, an OP pulse, which consists of two frequencies  both with $\sigma^-$-polarization, is applied  to bring the atom out of the $F=1$ manifold and back to  $|2, -2\rangle$ via spontaneous emission (Fig.~\ref{f-setup}A).
%brings the atom back to $|F=2, m_F=-2\rangle$ via spontaneous emission.
%, which consists of two frequencies with $\sigma^-$-polarization for the  to bring the atom out of $F=1$ manifold and back to  $|F=2, m_F=-2\rangle$ via spontaneous emission.
It is important that the optical pumping step scatter as few photons as possible to minimize recoil heating. %The OP consists of driving a $|F=1, m_F=-1\rangle$ into the $F=2$ manifold via the transition of $3P_{3/2} D2 line$
Therefore, to optically pump atoms from $|2,-1\rangle$ into $|2, -2\rangle$ while keeping the target state dark, we apply light resonant with  $3^2P_{1/2}$~\cite{Monroe1995, Grobner2017}~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3^2P_{3/2}$, from which the $|2, -2>$ state could always scatter a photon via the excited $|F'=3, m_F'=-3>$ state.}.
%~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3P_{3/2}$, from which the $|2, -2 \rangle$}

\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping motional-state redistribution and  Raman coupling for large  LD parameters for the axial direction ($z$). The range plotted covers $95$\% of the initial thermal distribution.
   % The range plotted covers $95$\% of the initial thermal distribution.
    (A) Motional state distribution %averaged over photon emission pattern
    after one OP cycle
    for different initial states motion, $n_{\textrm{init}}$. Due to photon-recoil and the large LD parameter, $\eta^{OP}_z=0.6$,
    there is a high  probability of $n$ changing.         (B) Matrix elements for Raman transition deviate from
    $\sqrt{n}$ scaling with multiple minima. During cooling, we utilize the fact that high motional states couple most effectively to sidebands with large $|\Delta n|$.
    \label{f-ld}}
\end{figure}

%During optical pumping, for an atom starting in motional level $n_{\textrm{init}}$, the probability of recoil to another state $n$ is determined by the squared matrix element $|\langle n_{\textrm{init}}|e^{i \Delta k z}| n\rangle|^2$ and causes motional-state redistribution as shown in Fig.~\ref{f-ld}A.  Here $\Delta k$ is the wave vector difference between absorbed and emitted photons. The LD parameters for OP are $\{\eta^{\textrm{OP}}_x,\eta^{\textrm{OP}}_y,\eta^{\textrm{OP}}_z\} = \{0.241(2), 0.206(1), 0.602(5)\}$.  As noted already, prior to RSC, the weak axial direction %, $(\eta^{\textrm{OP}}_{z,\textrm{eff}})^2>1$ is far outside of the LD regime, and the probability of recoil per OP step is large, possibly leading to runaway heating.

%Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
%The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$,
%where the wave vector is replaced by the wave vector difference $\Delta k$
%between the two beams that drive the Raman transition.
%The Raman transitions in our configuration have LD parameters of
%$\{\eta^R_x,\eta^R_y,\eta^R_z\} = \{0.341(2), 0.291(1), 0.40(1)\}$.
%To offset heating from OP initially when $n$ is large, higher-order Raman  sidebands ($\Delta n <-1$)   remove several motional quanta in a single cooling pulse (see Fig.~\ref{f-ld}B).
%Because the coupling strengths of different orders do not reach minima for the same state of motion $n$,
%using multiple orders of sidebands avoids accumulations of population
%near the coupling minima.

During optical pumping, photon recoil can lead to significant heating, especially along the weakly-confined axial direction ($z$).  For an atom starting in an axial motional level $n_{\textrm{init}}$, the probability to end in another axial state $n$ after absorbing and re-emitting one photon is determined by an angular integral of the squared matrix element $|\langle n_{\textrm{init}}|e^{i \Delta k_{OP} z}| n\rangle|^2$ \cite{ItanoWineland1979}. Here $\Delta k_{OP}$ is the wave vector difference between absorbed and emitted photons, projected onto the $z$-axis.  As shown in Fig.~\ref{f-ld}A, the probability of gaining motional energy during OP grows with $n_{\textrm{init}}$, and is an important effect for sideband-cooling outside the Lamb-Dicke regime.

Fortunately, the large LD parameters in our system also provide opportunities to overcome OP heating.
The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$,
where $\Delta k$ is the wave vector difference
between the two beams that drive the Raman transition.  Here $z_0=\sqrt{\hbar/(2 m \omega)}$ is the ground state wavefunction spread ($m$ is the atomic mass).
The Raman transitions in our configuration have LD parameters of
$\{\eta^R_x,\eta^R_y,\eta^R_z\} = \{0.341(2), 0.291(1), 0.40(1)\}$.
To offset heating from OP initially when $n$ is large, higher-order Raman  sidebands ($|\Delta n| > 1$)   remove several motional quanta in a single cooling pulse (see Fig.~\ref{f-ld}B).
Because the coupling strengths of different orders do not reach minima for the same state of motion $n$,
using multiple orders of sidebands avoids accumulations of population
near the coupling minima.

Taking the large motional-state changing heating and cooling sources
 into account,
it is not immediately clear that ground-state cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search~\cite{Dalibard1992}.
%and to find a robust cooling sequence~\footnote{See supplemental material}.
In particular, we find that alternating the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and $\Delta n=-2$ and $\Delta n=-1$ for the radial directions
eliminates the accumulation of population in motional states that have zero Raman coupling, which would halt the cooling process.
The simulation also indicates that setting the coupling strength of each  sideband
to drive a Rabi $\pi$-pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in Fig.~\ref{f-ld}B)  yields efficient cooling, initially.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state, so the final  cycles utilize only the $\Delta n=-1$ sideband while alternating between the three  axes.

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed  sideband orders ($\Delta n=-8$ and $\Delta n=-7$)
and decreasing the orders to the next pair after every $6$ to $15$ cycles.
This process is repeated until  $\Delta n=-1$, and most of the
population is in the first few excited states. We then switch to cooling only on $\Delta n=-1$ with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
Radial cooling is performed similarly to  axial cooling with the initial cooling of $\Delta n=-2$ and $\Delta n=-1$ and switching to $\Delta n=-1$ only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which is then used to calibrate experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates.

There are two additional important challenges in cooling single sodium atoms.
First, the  initial temperature populates high motional states,
causing the atoms to sample the anharmonicity of the trap away from the center. Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.
In the paraxial approximation, we find $A_{i,n}=\frac{-3n\hbar}{4\pi m d_i^2}$, where $d_i$ equals the beam radius for the radial directions and $d_z\approx\pi w_{0,x}w_{0,y}/\lambda_{\textrm{trap}}$.
%We find $A_{i,n}=\frac{3n \hbar}{4 \pi m d_i^2}$  where $d_i=w_{0,x},w_{0,y}$ for the radial directions and $d_z\approx \pi w_{0,x}w_{0,y}/\lambda$ for the axial direction.
Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{-1.1, -2.0, -0.16\}n$~kHz. For mixed states, this  broadens high-order sidebands due to the $n$-dependence of the transitions.
In order to mitigate this, we drive the sidebands with a large  Rabi frequency
and short pulses to Fourier broaden the spectrum. %, and use Blackman pulse envelopes to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}.
However, anharmonicity  may still shift high-$n$ states out of resonance and prevent effective cooling.
%Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.  We find $A_{i,n}=\frac{3n \hbar}{2 m d_i^2}$  where $d_i=w_0$ for the radial directions and $d_z=\pi w_0^2/\lambda$  Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{1.3, 1.2, 0.2\}n$~kHz.  Anharmonicity broadens the high-order axial transitions and may shift high-$n$ states out of resonance to prevent effective cooling.

@@

We measure a heating rate that corresponds to  decreasing  3D ground state population at a rate  of  $\sim0.9$\%/ms.
The  rate is consistent with off-resonant scattering of the trapping light~\cite{Grimm2000}, and is predominantly in the axial direction where the trapping beam propagates.
%The ground state probability achieved is currently limited by off-resonant scattering from the Raman beams, which are measured to be between $3$ to $15$ kHz for the detuning and power we use, and the resonance shift caused by magnetic field fluctuations.
